---
title: "Covid91 vaccine study Final2021"
author: "Kimberly Orantes"
date: "November 23, 2021"
output: 
    html_document:
    toc: true
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))

FinalData <- FinalData %>% mutate_if(is.character, as.factor)
FinalData$treatment <- relevel(FinalData$treatment, ref = "placebo")
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Guys

```{r}
Guys <- filter(FinalData, sex=="guy")
```

# Graphical Results
```{r}
barchartGC(~infected + treatment,data=Guys)
barchartGC(~infected + treatment,data=Guys, type="percent")
```

The frequency bar chart compares the frequency of guys who received the study drug Covid91 and those that received the placebo when they have covid or are OK. The frequency bar chart show that the majority population of the guys subgroup were OK. The percent bar chart compares the percentages of patients received the study drug Covid91 and those that receive the placebo. The percent bar chart shows that about 10% more guys that received the study drug Covid91 were infected with covid than the guys that received the placebo. This shows that the study drug Covid91 is ineffective for guys and it increases the likelihood of becoming infected with covid rather than decreasing the odds.       
   
# Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Guys)
rowPerc(table1)
colPerc(table1)
```

The first table shows that 58 people who received the study drug Covid91 got infected with covid and 42 people who received the placebo got infected with covid. The second table shows that a majority of the patients were OK, 97.5% received the placebo and were OK but 2.5% got infected with covid. 96.3% were OK when they received the study drug Covid91 but 3.7% got infected with covid after receving the drug. This shows that the drug increases the likelihood of contracting covid.         
    
# Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Guymod <- fisher.test(table1)
guyeff <- (Guymod$estimate/(Guymod$estimate+1))*100
Guyeffupper <- Guymod$conf.int[2]
Guyeffuppereff <- (Guymod$conf.int[2]/(Guymod$conf.int[2]+1))*100
Guyefflower <- Guymod$conf.int[1]
Guyefflowereff <- (Guymod$conf.int[1]/(Guymod$conf.int[1]+1))*100
options(digits=3)
```

The hypothesis of this study is that there is a noticeable difference between those that received the study drug Covid91 and those that received a placebo in getting infected with COVID. The null hypothesis would then be that there is little to no difference between those that received the study drug Covid91 and those that received a placebo in getting infected with COVID. The Chi-squared  and Fisher tests both produced a p-value less than 0.05 which is the probability under the null hypothesis conditions. The Fisher test calculated the odds ratio of guys getting infected with COVID when taking the study drug Covid91 was `r Guymod$estimate` times more than those who received the placebo. This meant the guys were `r Guymod$estimate` times more likely in getting covid because of the study drug Covid91. Based on these results, we reject the null hypothesis because guys are more likely to get infected with COVID after receiving the study drug Covid91. The efficacy of the drug for the guys is `r guyeff`% with a 95% confidence interval of `r Guyefflowereff` to `r Guyeffuppereff`. From these results, I would advise guys against receiving the Covid91 study drug.  
   
# Gals

```{r}
Gals <- filter(FinalData, sex=="gal")
```

# Graphical Results
```{r}
barchartGC(~infected + treatment,data=Gals)
barchartGC(~infected + treatment,data=Gals, type="percent")
```

The frequency bar chart compares the frequency of gals who received the study drug Covid91 and those that received the placebo when they have covid or are OK. The frequency bar chart show that the majority population of the gals subgroup were OK. The percent bar chart compares the percentages of patients received the study drug Covid91 and those that receive the placebo. The percent bar chart shows that about 30% more gals that received the placebo were infected with covid than the gals who received the study drug Covid91. This shows that the study drug Covid91 is effective for gals since it decreased the likelihood of getting infected with covid.       

# Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=Gals)
rowPerc(table1)
colPerc(table1)
```

The first table shows that 67 gals who received the placebo got infected with covid while only 33 gals who received the study drug Covid91 got infected with covid. The second table shows that the gals that received the placebo 8.7% of them got covid and the gals that received the drug only 4.4% of them got covid. These results show that the study drug Covid91 is effective for gals since it decreases the likelihood of getting infected with covid.  
    
# Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Galmod <- fisher.test(table1)
galeff <- (Galmod$estimate/(Galmod$estimate+1))*100
Galeffupper <- Galmod$conf.int[2]
Galeffuppereff <- (Galmod$conf.int[2]/(Galmod$conf.int[2]+1))*100
Galefflower <- Galmod$conf.int[1]
Galefflowereff <- (Galmod$conf.int[1]/(Galmod$conf.int[1]+1))*100
options(digits=3)
```

The hypothesis of this study is that there is a noticeable difference between those that received the study drug Covid91 and those that received a placebo in getting infected with COVID. The null hypothesis would then be that there is little to no difference between those that received the study drug Covid91 and those that received a placebo in getting infected with COVID. The Chi-squared  and Fisher tests both produced a p-value less than 0.05 which is the probability under the null hypothesis conditions. The Fisher test calculated the odds ratio of gals getting infected with COVID when taking the study drug Covid91 was `r Galmod$estimate` times less than those who received the placebo. This meant the guys were `r Galmod$estimate` times less likely to get covid since they received the study drug Covid91. Based on these results, we reject the null hypothesis because gals are less likely to get infected with COVID after receiving the study drug Covid91. The efficacy of the drug for the gals is `r galeff`% with a 95% confidence interval of `r Galefflowereff` to `r Galeffuppereff`. From these results, I would advise gals to receive the Covid91 study drug.


# LGBTQ


# Drug Users

# Discussion and Conclusion
