---
title: "Covid91 vaccine study Final2021"
author: "Kimberly Orantes"
date: "November 23, 2021"
output: 
    html_document:
    toc: true
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))

FinalData <- FinalData %>% mutate_if(is.character, as.factor) 
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Guys

```{r}
Guys <- filter(FinalData, sex=="guy")
```

# Plot
```{r}
barchartGC(~infected + treatment,data=Guys)
barchartGC(~infected + treatment,data=Guys, type="percent")

```

The frequency barchart show that the majority population of the guys subgroup were not infected. The percent barchart shows that there isn't much difference in the effectiveness of the drug for covid because it was close to the percentages of the placebo for both the infected and not infected guys.    
   
# Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Guys)
rowPerc(table1)
colPerc(table1)
```

Those that were infected with covid, only 58.16 % of them received the drug and only 41.84% received the placebo. The numerical results show that majority of the guys were ok and uninfected.     
    
# Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-square test shows that there is a significant difference in the expected counts made by the null hypothesis and the observed counts because the p-value is less than 0.05. The Fisher test produced a p-value less than 0.05 as well. The Fisher test also calculated the odds ratio of guys getting infected with COVID when taking the drug was 1.47. The 95% confidence interval is from 1.18 to 1.85. With the results of these tests we can conclude that the drug has a significant effect on the likelihood of guys becoming or not becoming infected with COVID therefore we can reject the null hypothesis.   
   
# Gals


# LGBTQ


# Drug Users

# Discussion and Conclusion
